Measurement and Probability in Fuzzy Quantum Theories

### 2016-12-10 prop.tex
prop.tex \section{Introduction} Reorginize to make more like a paper...

This is a \emph{theoretical} investigation of \emph{experimental}
physics using \emph{computational} methods. All experiments and computations
are processes bounded in space, time, energy, and other resources~\cite{Jaeger2007,Li2009}.
Yet, for centuries, the mathematical formalization of such processes
has been founded on the infinitely precise real or complex numbers~\cite{Kent1999,CliftonKent2000}.
Indeed, almost every description of quantum mechanics, quantum computation,
or quantum experiments refers to entities such as $\rme$, $\pi$,
$\sqrt{2}$, etc. From a computational perspective, such numbers do
not exist in their entirety ``for free.'' For example, the state
of the art algorithms for computing the $n$th binary digit of~$\pi$
require on the order of $O(n\log^{O(1)}(n))$ operations~\cite{journals/moc/BaileyBP97}.
In other words, simply referring to the $n$th digit of $\pi$ requires
more and more resources as $n$ gets larger. Taking such resource
bounds into consideration is what founded computer science as a discipline
and is crucial for understanding the very nature of computation and,
following Feynman~\cite{Feynman1982Simulating}, Landauer~\cite{Landauer1996188},
and others, for understanding the very nature of physical processes.

We propose to revisit quantum mechanics, quantum information, and
quantum computation from this resource-aware perspective. Our initial
results in that domain showed how subtle the issues can
be~\cite{usat,geometry2013,DQT2014}: a straightforward replacement of
the complex numbers by a finite field yields a variant of quantum
mechanics in which computationally hard problems like UNIQUE-SAT
(which decides whether a given Boolean formula is unsatisfiable or has
exactly one satisfying assignment) can be deterministically solved in
constant time. To eliminate such unrealistic theories requires
delicate analysis of the structure of the Hilbert space, the process
of observation, and the notion of probability teasing apart their
reliance on the infinitely precise real
numbers~\cite{geometry2013,DQT2014}. In this proposal, our aim is to
shift focus from the infinitely-specified but not directly observable
quantum states, to observable measurable properties of quantum systems
and their probabilities. Furthermore, we insist that our theories of
measurement and probability only refer to finitely communicable
evidence within feasible computational bounds. It follows that states,
observations, and probabilities all become ``fuzzy'', i.e., specified
by intervals of confidence that can only increase in precision if the
available resources increase proportionally. Our notion of ``fuzzy
quantum mechanics'' is related to existing
work~\cite{GranikCaulfield1996,Pykacz2013,SNL2009,Gudder2005,aerts1993physical}
but, as will be explained in more detail, is distinguished by its
unique computational character.

@@

In particular, the objectives to be achieved include:
\begin{itemize}
\item A computational framework for quantum measurement where quantum
  systems are described in terms of interactive processes. The
  framework takes care of the resources required (including the ones
  related to the observer) to determine a given property of the
  measured system.
\item Develop a measurement framework based on quantum interval-valued
  probabilities measures.
\item Assess the validity of fundamental theorems of quantum
  mechanics, such as Gleason, Bell, and Kochen-Specker, in a quantum
  formulation stripped away of infinite resources. We will aim to
  provide a definite answer to the Meyer-Mermin
  debate~\cite{PhysRevLett.83.3751, Mermin1999, BarrettKent2004} on
  the impact of finite precision measurements on the relevance of the
  Kochen-Specker
  theorem~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}.
\item Study the computational complexity of quantum theories and
  simulations based on discrete (computable) fields.
\item Study the problem of finding a complete set of mutually unbiased
  bases
  (SIC-POVM)~\cite{1751-8121-45-8-085306,1751-8121-47-33-335302,Caves2002}
  over finite Galois fields.
\end{itemize}
We argue that no matter what the results are their impact will be
strong. Positive results that formulate a computable, with clear
resource bounds, theory of quantum measurement and quantum probability
would be essential for understanding and realizing quantum
computation. Any negative results would redirect research that aims at
realizing quantum computation to other approaches. At an intellectual
level, the results have the potential to solve or clarify several
paradoxes in quantum mechanics, quantum information, computability,
complexity theory, and perhaps quantum gravity.

>>>>

This is a \emph{theoretical} investigation of \emph{experimental}
physics using \emph{computational} methods. All experiments and computations
are processes bounded in space, time, energy, and other resources~\cite{Jaeger2007,Li2009}.
Yet, for centuries, the mathematical formalization of such processes
has been founded on the infinitely precise real or complex numbers.
Our purpose here is exploit the consequences of replacing infinitely-specified
quantum probabilities by finite number of intervals used in interval-valued
probability measures; in particular, we show that a quantum interval-valued
probability measure may not be induced by a infinitely-precise state,
but it might be more and more likely to be identified to a particular
state when the measurement resources increase.

Indeed, almost every description of quantum mechanics, quantum computation,
or quantum experiments refers to entities such as $\rme$, $\pi$,
$\sqrt{2}$, etc. From a computational perspective, such numbers do
not exist in their entirety ``for free~\cite{Kent1999,CliftonKent2000}.''
For example, the state of the art algorithms for computing the $n$th
binary digit of~$\pi$ require on the order of $O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, simply
referring to the $n$th digit of $\pi$ requires more and more resources
as $n$ gets larger. Taking such resource bounds into consideration
is what founded computer science as a discipline and is crucial for
understanding the very nature of computation and, following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, for understanding the
very nature of physical processes.

We propose to revisit quantum mechanics, quantum information, and
quantum computation from this resource-aware perspective. Our initial
results in that domain showed how subtle the issues can
be~\cite{usat,geometry2013,DQT2014}: a straightforward replacement of
the complex numbers by a finite field yields a variant of quantum
mechanics in which computationally hard problems like UNIQUE-SAT
(which decides whether a given Boolean formula is unsatisfiable or has
exactly one satisfying assignment) can be deterministically solved in
constant time. To eliminate such unrealistic theories requires
delicate analysis of the structure of the Hilbert space, the process
of observation, and the notion of probability teasing apart their
reliance on the infinitely precise real
numbers~\cite{geometry2013,DQT2014}.

In this proposal, our aim is to
shift focus from the infinitely-specified but not directly observable
quantum states, to observable measurable properties of quantum systems
and their probabilities. Furthermore, we insist that our theories of
measurement and probability only refer to finitely communicable
evidence within feasible computational bounds. It follows that states,
observations, and probabilities all become ``fuzzy'', i.e., specified
by intervals of confidence that can only increase in precision if the
available resources increase proportionally. Our notion of ``fuzzy
quantum mechanics'' is related to existing
work~\cite{GranikCaulfield1996,Pykacz2013,SNL2009,Gudder2005,aerts1993physical}
but, as will be explained in more detail, is distinguished by its
unique computational character.

@@

In particular, the objectives to be achieved include:
\begin{itemize}
\item Develop a measurement framework based on quantum interval-valued
  probabilities measures.
\item Assess the validity of fundamental theorems of quantum
  mechanics, such as Gleason, Bell, and Kochen-Specker, in a quantum
  formulation stripped away of infinite resources. We will aim to
  provide a definite answer to the Meyer-Mermin
  debate~\cite{PhysRevLett.83.3751, Mermin1999, BarrettKent2004} on
  the impact of finite precision measurements on the relevance of the
  Kochen-Specker
  theorem~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}.
\end{itemize}
